\documentclass[]{ccs-thesis}
% options:
% [germanthesis] - Thesis is written in German
% [plainunnumbered] - Don't print numbers on plain pages
% [earlydraft] - Settings for quick draft printouts
% [watermark] - Print current time/date at bottom of each page
% [phdthesis] - switch to PhD thesis style
% [twoside] - double sided
% [cutmargins] - text body fills complete page

% Required package
\usepackage{tikz}
\usepackage{outlines}
\usepackage[underline=true,rounded corners=false]{pgf-umlsd}
% Protocol Header file
\usepackage{bytefield}
\usepackage{makecell}
\usepackage{tabularx}
\usepackage{tabularray}
% table coloring
\usepackage{colortbl}
\usepackage{lipsum}
% block diagram
\usepackage{amsmath}
\usetikzlibrary{shapes.geometric,positioning}
% listings without linebreak
\usepackage{adjustbox}


\lstdefinestyle{customc}{
  belowcaptionskip=1\baselineskip,
  breaklines=true,
  frame=single,
  xleftmargin=\parindent,
  language=C++,
  showstringspaces=false,
  basicstyle=\footnotesize\ttfamily,
  keywordstyle=\bfseries\color{green!40!black},
  commentstyle=\itshape\color{purple!40!black},
  identifierstyle=\color{blue},
  stringstyle=\color{orange},
}

\lstset{escapechar=@,style=customc}

% Choose one of the following lines.
%\group{Cooperative Mobile Systems}
\group{Distributed Embedded Systems}

% Author name. Separate multiple authors with commas.
\author{Maximilian W. Gotthardt}
\birthday{13. July 1993}
\birthplace{Berlin}

% Title of your thesis.
\title{Evaluation of protocols \\for control stage lighting}

% Degree program.
\thesistype{Bachelor Thesis in Computer Engineering}\thesiscite{Bachelor Thesis~(Bachelorarbeit)}

% List of advisors, separated by commas. Delete if identical to referees or inapplicable.
\advisors{Anatolij Zubow}

% List of referees, separated by commas.
\referees{Falko Dressler, Thomas Sikora}

% Abbreviations own
\acrodef{RR}{Rapid Repetition}
\acrodef{MCU}{Microcontroller Unit}

% Abbreviations light protocols
\acrodef{DMX}{Digital Multiplex}
\acrodef{CANhashing}[CAN]{Content Addressable Network}{extra={when referring to the distributed hash table}}
\acrodef{CANproto}[CAN]{Controller Area Network}{extra={when referring to the bus protocol}}

% Abbreviations Networking
\acrodef{IEEE}{Institut of Electrical and Electronics Engineers}
\acrodef{OSI}{Open Systems Interconnection}
\acrodef{MSDU}{MAC Service Data Unit}
\acrodef{DL}{Data Link Layer}
\acrodef{LLC}{Logic Link Control}
\acrodef{MAC}{Media Access Control}
\acrodef{IP}{Internet Protocol}
\acrodef{CD}{Collision Detection}
\acrodef{CSMA/CA}{Carrier-sense Multiple Access with Collision Avoidance}
\acrodef{DCF}{Distributed Coordination Function}
\acrodef{PCF}{Point Coordination Function}
\acrodef{DIFS}{DCF Inter Frame Spaces}
\acrodef{SIFS}{Short Inter Frame Spaces}
\acrodef{NAV}{Network Allocation Vector}
\acrodef{PHY}{Physical Network Layer}
\acrodef{CW}{Contention Window}
\acrodef{MPDU}{Mac Protocol Data Unit}
\acrodef{DSSS}{Direct Sequence Spread Spectrum}
\acrodef{FHSS}{Frequency Hopping Spread Spectrum}
\acrodef{STA}{Station}
\acrodef{AP}{Access Point}
\acrodef{WES}{Wireless Endsystem}
\acrodef{BSS}{Basic Service Set}
\acrodef{OFDM}{Orthogonal Frequency Division Multiplexing}
\acrodef{MIMO}{Multiple Input-Multiple Output}
\acrodef{WLAN}{Wireless Local Area Network}
\acrodef{LAN}{Local Area Network}
\acrodef{BSSID}{Basic Service Set Identifier}
\acrodef{UDP}{User Datagram Protocol}

% Abbreviations other
\acrodef{UART}{Universal Asynchronous Receiver Transmitter}

\begin{document}

\pagenumbering{roman}

\maketitle

\thispagestyle{empty}

\cleardoublepage

\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
\begin{otherlanguage*}{american}

% about 1/2 page:
% \begin{enumerate}
% 	\item Motivation (Why do I care?)
% 	\item Problem statement (What problem are I trying to solve?)
% 	\item Approach (How did I go about it)
% 	\item Results (What's the answer?)
% 	\item Conclusion (What are the implications of the answer?)
% \end{enumerate}

In the field of lighting and stage technology, the challenge of controlling the individual installations 
quickly and without complications is a recurring one.
Established solutions are realized via cables ensuring no interference,
while wireless solutions are more flexible.
Crowded events such as concerts may cause interference in the radio channel, 
resulting in a drop in reliability.

While controlling many nodes with only small payloads, the overhead is significant.
This thesis targets the problem of efficient and reliable wireless control of light installations 
by proposing a low-overhead data link layer solution on top of IEEE 802.11 physical layer
supporting both reliable unicast and broadcast operations.
The latter is scalable which is important in case a large group of nodes needs to be controlled.

The proposed protocols were analyzed analytically 
as well as by means of experiments in a small indoor testbed using the ESP-NOW protocol. 
Results reveal that the proposed approaches are able to outperform state-of-the-art solutions 
(i.e., ARTNet over TCP/UDP and Wi-Fi stack) through the use of flatter protocol stacks and efficient broadcasts.
By applying unsolicited delayed repetitions in the application layer 
the success ratio can be raised to over 99\%, even with weak connections.
Moreover, it enables time synchronized operation on a set of nodes, e.g., 
for the purpose of time controlled configuration of lighting, 
which is not performed with state-of-the-art wireless solutions.

\end{otherlanguage*}

\chapter*{Kurzfassung}
\addcontentsline{toc}{chapter}{Kurzfassung}
\begin{otherlanguage*}{ngerman}

Im Bereich der Licht- und Bühnentechnik ist es herausfordernd, 
die verteilten Lichtanlagen schnell und unkompliziert anzusteuern.
Etablierte Lösungen werden über Kabel realisiert, die keine Störungen verursachen,
während aber drahtlose Lösungen deutlich flexibler sind.
Insbesondere bei Großveranstaltungen wie Konzerten kann es zu Störungen im Funkkanal kommen,
was zu einer Abnahme der Zuverlässigkeit führt.

Diese Arbeit befasst sich mit dem Problem der effizienten und zuverlässigen drahtlosen Steuerung von Lichtinstallationen.
Durch die Steuerung vieler Knoten mit geringen Nutzlasten fällt der Overhead erheblich ins Gewicht.
Lösung mit geringem Overhead auf der Datenübertragungsschicht des IEEE 802.11 werden eingeführt,
die sowohl zuverlässigen Unicast- als auch Broadcast-Betrieb unterstützten.
Letzterer ist skalierbar, was wichtig ist, wenn eine große Gruppe von Knoten angesteuert wird.

Die vorgeschlagenen Protokolle wurden analytisch analysiert,
sowie anhand von Experimenten in einer kleinen Indoor-Testumgebung unter Verwendung des ESP-NOW-Protokolls getestet.
Die Ergebnisse zeigen, dass die vorgeschlagenen Ansätze in der Lage sind, 
den Stand der Technik (d.h. ARTNet über TCP/UDP und Wi-Fi Stack),
durch die Verwendung von flacheren Protokollstapeln und effizienten Übertragungen, zu übertreffen.
Durch den Einsatz von unaufgeforderten verzögerten Wiederholungen in der Anwendungsschicht,
kann die Erfolgsquote selbst bei schlechtem Empfang auf über 99 \% gesteigert werden.
Darüber hinaus ermöglichen die Protokolle einen zeitsynchronen Betrieb über alle Knoten hinweg,
was insbesondere beim parallelen Schalten der Lichter wichtig ist 
und bei kabellosen modernen Funklösungen nur eingeschränkt möglich ist. 

\end{otherlanguage*}
\acresetall

\cleardoublepage
\tableofcontents

\cleardoublepage
\pagenumbering{arabic}


\include{thesisIntro}

\include{thesisRelated}

\include{thesisFund}

\include{thesisProposed}

\include{thesisEvaluation}

% \include{Guidelines}

\cleardoublepage

\listofabbreviations
\clearpage

\listoffigures
\clearpage

\listoftables
\clearpage

\lstlistoflistings
\clearpage

\printbibliography

\end{document}
